Automatic extraction of symbolic network signatures from malware code

### 2016-10-30 paper.tex
Added progress milestone 1

\documentclass[]{article}

\usepackage{hyperref}

@@

\section{Project Proposal}
\textbf{Tasks}:
\begin{itemize}
	\item Identify and characterize features of common malware code - 1-2 weeks
	\item Develop the abstract semantics and operators for extracting  
	symbolic network signatures from malware code - 2-3 weeks
	\item Modifying/Extending existing tools like Codesurfer for symbolic 
	network signature extraction - 2-3 weeks
\end{itemize}

@@

\noindent\textbf{Evaluation}: 
The primary metrics of evaluation are time taken for analysis, accuracy and 
generality of the symbolic signatures. Performance analysis can be performed 
with different malware programs with varying lines of code to understand
the time complexity of the analysis. The quality of the symbolic network
signatures is a tougher metric to measure. Ideally, the symbolic network 
signature must only include packets which will make the program reach the
specified point. If the signature is more general, legitimate packets may 
be considered as malicious and blocked. On the other hand, the signature
may miss actual malware traffic because the signature does not encompass
all inputs leading to the point. We would try to address these trade-offs in
our evaluation.
\bibliographystyle{abbrv}
\bibliography{references}

>>>>

\documentclass[]{article}
\usepackage[xcolor]{changebar} 
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{listings}
\newcommand{\deleted}[1]{%
	\cbcolor{red}
	\begin{changebar}
		#1
	\end{changebar}%
}%

\newcommand{\added}[1]{%
	\cbcolor{green}
	\begin{changebar}
		#1
	\end{changebar}%
}%
\usepackage{tikz}
\usetikzlibrary{automata,positioning,arrows}

@@

\section{Project Proposal}
\textbf{Tasks}:
\begin{itemize}
	\item Identify and characterize features of common malware code - 1-2 weeks
	\added{Refer to ---}
	\item \deleted{Develop the abstract semantics and operators for extracting  
	symbolic network signatures from malware code - 2-3 weeks} \\
	\added{Automatically generate a symbolic automaton of the given program
		based on packet and variable state - 2-3 weeks}
	\item Modifying/Extending existing tools like Codesurfer for symbolic 
	network signature extraction - 2-3 weeks
\end{itemize}

@@

\noindent\textbf{Evaluation}: 
The primary metrics of evaluation are time taken for analysis, accuracy and 
generality of the symbolic signatures. Performance analysis can be performed 
with different malware programs with varying lines of code to understand
the time complexity of the analysis. The quality of the symbolic network
signatures is a tougher metric to measure. Ideally, the symbolic network 
signature must only include packets which will make the program reach the
specified point. If the signature is more general, legitimate packets may 
be considered as malicious and blocked. On the other hand, the signature
may miss actual malware traffic because the signature does not encompass
all inputs leading to the point. We would try to address these trade-offs in
our evaluation.

\appendix
\section{Malware program model}
The malware program listens to incoming packets, and the
network signature is the set of incoming packets which triggers
an attack by the program. We propose a reasonable abstraction
of the code of malware programs.

We model a malware program as a infinite reactive system 
of the form:
\begin{lstlisting}	
while(True) {
	packet = socket.receive();
	...
}
\end{lstlisting}
Each iteration of the loop, the program receives a
packet, and based on current state and packet received,
it performs certain actions. Using static analysis
techniques, we can identify two set of variables, \emph{state}
and \emph{packet} variables. State variables represent the 
program state in terms of the protocol, which packet variables
are used to represent the current packet being processed by
the program.
\begin{lstlisting}	
while(True) {
	packet = socket.receive();
	...
	if (a == 1 and packet[8] = "b") {
		a = a + 1;
	}
	...
}
\end{lstlisting}	
In the above code snippet, $a$ is a state variable
and $packet$ is a packet variable. This also
illustrates the packet processing 
function of the program, which based on the current
input packet and state variables, either transitions
to a new program state or performs an action (attack or
send packets). 
\begin{lstlisting}	
a = 0;
while(True) {
packet = socket.receive();
	if (a == 0 and packet[0] = "a") {
		a = a + 1;
	}
	if (a == 1 and packet[0] = "b") {
		a = a + 1;
	}
	if (a == 2 and packet[0] = "c") {
		attack()  [the attack point is provided as input to analysis]
	}	
}
\end{lstlisting}	
The above code snippet represents the complete model
of the malware program, where based on input packets and
current state, the program transitions to a new state. 
The symbolic automaton for this program is 
shown in \Cref{fig:progautomaton}. Network signatures
for the malware are all paths from start state ($a=0$)
to state of attack. In this case, the sequence of packets
in order to initiate an attack is $a* \rightarrow b* \rightarrow c*$.
\begin{figure}[h]
	\centering
	\begin{tikzpicture}[shorten >=0.5pt,node distance=3cm,on grid,auto] 
	\node[state, initial] (v0)   {$a = 0$}; 
	\node[state] (v1) [right=of v0] {$a = 1$}; 
	\node[state] (v2) [right=of v1] {$a = 2$};
	\node[state] (v3) [right=of v2] {$attack$}; 	
	\path[->] 
	(v0) edge node {$a*$} (v1)
	(v1) edge node {$b*$} (v2)
	(v2) edge node {$c*$} (v3);
	\end{tikzpicture}
	\caption{Symbolic automaton for malware program.}
	\label{fig:progautomaton}
\end{figure}
\bibliographystyle{abbrv}
\bibliography{references}
